---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sleepIPD

<!-- badges: start -->
<!-- badges: end -->

sleepIPD is an opinionated wrapper around the [GGIR](https://CRAN.R-project.org/package=GGIR) accelerometer processing scripts. The wrapper ensures that contributors can reprocess data in precisely the same way, reducing variations in the data that arise from different processing decisions. In addition to the reprocessing functions, sleepIPD also provides a series of helper functions to manage the reprocessing.

## Installation

This package is not available on CRAN. You can install from [GitHub](https://github.com/Motivation-and-Behaviour/sleepIPD) with:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("Motivation-and-Behaviour/sleepIPD")
```
## Documentation

You can find the documentation and getting started guide for sleepIPD [here](https://motivation-and-behaviour.github.io/sleepIPD/articles/sleepIPD.html). This also includes the steps for reprocessing if your data are large or more complex.
